A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-11-05 paper/paper.tex
paper: polish first few paragraphs

\section{Introduction}
\subsection*{Fibered vs indexed presentation of semi-simplicial and semi-cubical sets}
A family of sets can commonly be represented in two ways: as a family properly speaking, indexed over a given set $S$, or as a set $T$ together with a map specifying for each element of the set $T$ on which index in $S$ it depends. In the first case, we call it an \emph{indexed} presentation. In the latter case, the set associated to a given index is the fiber, i.e. inverse image, of the index under the map, so we call it a \emph{fibered} presentation. The two presentations are equivalent and the equivalence can be phrased concisely in the language of homotopy type theory~\citep{hottbook} as a fibered/indexed correspondence\footnote{At a discussion on the Zulip channel for the HoTT community, alternative names fibration/family correspondence and unbundled/bundled correspondence were proposed, among others. In our experience, the opposition fibered/indexed is spontaneously understood by many listeners. In particular, it echoes to the Grothendieck construction of fibered categories from indexed categories. The most elementary instance of the correspondence, with $\Type$ instead of $\U$, that is $(\Sigma S: \Type. (S \rightarrow T)) \simeq (T \rightarrow \Type)$, is sometimes called ``Grothendieck construction for dummies'' (its proof requires univalence, see \cite{hottbook}).}
\begin{equation*}
  \mbox{(fibered)}\qquad(\Sigma S: \U. (S \rightarrow T)) ~\simeq~ (T \rightarrow \U)\qquad \mbox{(indexed)}
\end{equation*}
where $\U$ represents in homotopy type theory the subset of types with a unique proof of equality (within some universe).

A \emph{presheaf} on an category is a family of sets connected with maps indexed over the morphisms of the category. As such, it lives on the indexed side of the correspondence, opposing on the fibered side to \emph{discrete Grothendieck fibrations}~\cite{LoregianRiehl20}. However, there are situations where a presheaf $F$ can also be seen as living on the fibered side of the correspondence. Indeed, when the indexing category is \emph{direct}, i.e. with a downwards-well-founded collection of non-identity morphisms, as is the case of semi-simplicial and semi-cubical sets, each $F(X)$ serving as source of a map $F(X) \stackrel{F(f)}{\rightarrow} F(Y)$ can be turned into a family $F(X)$ indexed on $F(Y)$, and, since the maps are well-founded downward, this can be iterated from below, eventually producing a collection of families which we call an \emph{indexed presentation} of the presheaf over the direct category.

For instance, in the case of a semi-cubical set~\citep{grandis03,buchholtz17}, presented with $2n$ face maps from the set of $n$-cubes to the set of $(n-1)$-cubes, the standard presheaf approach prescribes a family of sets fibered the ones over the others (up to cubical faces identities). Formulated in type theory, it corresponds to:

>>>>

\section{Introduction}
\subsection*{Fibered vs indexed presentation of semi-simplicial and semi-cubical sets}
A family of sets can commonly be represented in two ways: as a family properly speaking, indexed over the elements of a given set $S$, or as a set $T$ together with a map from $T$ to $S$, which specifies for each element of set $T$ its dependency on $S$. In the former case, we call it an \emph{indexed} presentation. In the latter case, the set associated to a given element of $S$ is the fiber of this element, so we call it a \emph{fibered} presentation. The two presentations are equivalent and the equivalence can be phrased concisely in the language of homotopy type theory~\citep{hottbook} as the fibered/indexed correspondence\footnote{In an informal discussion, alternative nomenclatures were proposed: fibration/family correspondence and unbundled/bundled correspondence. The fibered/indexed nomenclature echoes the Grothendieck construction of fibered categories from indexed categories. The most elementary instance of the correspondence, with $\Type$ instead of $\U$, is sometimes called ``Grothendieck construction for dummies'' (its proof requires univalence, see \cite{hottbook}).}.
\begin{equation*}
  \mbox{(fibered)}\qquad(\Sigma T: \U. (T \rightarrow S)) ~\simeq~ (S \rightarrow \U)\qquad \mbox{(indexed)}
\end{equation*}
where $\U$ represents in homotopy type theory the subset of types within a given universe where equality of any two elements has at most one proof.

A \emph{presheaf} on an category is a family of sets connected with maps indexed over the morphisms of the category. As such, it lives on the indexed side of the correspondence, contrasting with the fibered side, where we have \emph{discrete Grothendieck fibrations}~\cite{LoregianRiehl20}. However, there are situations where a presheaf $F$ can also be seen as living on the fibered side of the correspondence. This happens when the indexing category is \emph{direct}, or has a downwards-well-founded collection of non-identity morphisms. This is the case in semi-simplicial and semi-cubical sets, where each $F(X)$ serving as source of a map $F(X) \stackrel{F(f)}{\rightarrow} F(Y)$ can be turned into a family $F(X)$ indexed on $F(Y)$. Since the maps are well-founded downward, this can be iterated from below, eventually producing a collection of families which we call an indexed presentation of the presheaf over the direct category.

For instance, in the case of a semi-cubical set~\citep{grandis03,buchholtz17} presented with $2n$ face maps from the set of $n$-cubes to the set of $(n-1)$-cubes, the standard presheaf approach prescribes a family of sets fibered the ones over the others (up to cubical faces identities). Formulated in type theory, it corresponds to:
